Wind Veer and Speed in Turbulent Ekman Flow - Part 1: Scaling analysis and universal profile model

### 2024-10-17 write1/manuscript.tex
first complete draft

%
%\begin{figure}
%  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
%  \caption{Total velocity and veering (in degrees) vs inner and outer height.
%    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
%\end{figure} 
%
\section{Conclusions}
%
\subsection{Summary}
We investigate the wind veer in Ekman flow. Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.  
% 
This formulation is consistent with the DNS data and also yields reasonable results 
at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
%
The classic formulation of the Ekman layer, employing the surface boundary condition, is replaced by a modified solution that can be obtained by Ekman's system of governing equations, but using different boundary conditions that are more appropriate of the actual situation encountered in the planetary boundary layer. 
% 
The three parameters that characterize this boundary condition are estimated based on DNS data. 
% 
\par
% 
To quantify the spanwise velocity component consistently across the boundary layer, 
we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
\[V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+),\]
that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
% 
This scaling is derived here based on scaling considerations, and it is in excellent agreement with the DNS data available.   
% 
In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model mentioned above (Sec.~\ref{sec:ekman}).
% 
\par
%
The velocity profiles are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}. The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer. 
% 
In the inner layer, the logarithmic profile is (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
% 
In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).

\subsection{Applications \& Outlook} 
\begin{itemize}
\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
\item
\end{itemize}

@@

\appendix
\section{Old Stuff}
\todo{OLD FROM HERE} 
Below this region, the gradients in span-wise velocity are rather small and the span-wise velocity
monotonically approaches its surface boundary condition $V(z=0)=0$.
%
While the streamwise velocity follows a universal inner scaling that has acquired its universal, $Re$-independent shape for $Re_D> \mathcal{O}\left(10^3\right)$, the span-wise component that defines how the velocity vector veers when the surface is approached,
does not collapse in inner units, and there is, most importantly no sign of convergence even at the highest Reynolds numbers for
which simulations where carried out.
%
Even though the simplest assumption $V=0$ is reasonable for the lower part of the surface surface layer ($z^-<10^{-3}$), it does
not appropriately capture the profile in the rest of the surface layer:

>>>>

%
%\begin{figure}
%  \includegraphics[width=\textwidth]{../plot/mag_dir.pdf} 
%  \caption{Total velocity and veering (in degrees) vs inner and outer height.
%    Dashed lines show DNS data, thick lines are from semi-empirical theory.} 
%\end{figure} 
%
\section{Conclusions}
%
We investigate the wind veer in Ekman flow. Based on scaling considerations and direct numerical simulation spanning on decade in external separation from $Re_\Lambda=1.25 \times 10^5$ to $Re_\Lambda=1.28\times 10^6$, we derive a universal formulation for the both horizontal components of the velocity profile in a stationary, smooth Ekman layer.  
% 
This formulation is consistent with the DNS data and also yields reasonable results 
at geophysical scale separation; the logarithmic law of the mean velocity is recovered with the well-known limits and deviations towards the surface and Ekman layer. 
%
The classic formulation of the Ekman layer, employing the surface boundary condition, is replaced by a modified solution that can be obtained by Ekman's system of governing equations, but using different boundary conditions that are more appropriate of the actual situation encountered in the planetary boundary layer. 
% 
The three parameters that characterize this boundary condition are estimated based on DNS data. 
% 
\par
% 
To quantify the spanwise velocity component consistently across the boundary layer, 
we derive a universal scaling of the spanwise velocity component in a shear-aligned reference frame. For the \emph{inner layer}, we find the mixed scaling  
\[V^{\alpha_\star}/G = \frac{1}{\delta^+} f(z^+),\]
that is, the spanwise velocity normalized by the outer velocity scale is a universal function of the inner height and Friction Reynolds number $Re_\tau = \delta^+$. 
% 
This scaling is derived here based on scaling considerations, and it is in excellent agreement with the DNS data available.   
% 
In the outer layer, the spanwise velocity follows outer scaling, consistently with the Ekman model mentioned above (Sec.~\ref{sec:ekman}).
% 
\par
%
The velocity profiles are shown in inner and outer scaling as well as in ho\-do\-graph- view in Fig.~\ref{fig:summary}. The shear-aligned velocity deficit is shown in outer scaling highlighting the universality in the outer layer. 
% 
In the inner layer, the logarithmic profile (blue, panel \textbf{(b)}) is ecnountered across a widening region for increasing Reynolds number. 
% 
In this simple inner scaling, the spanwise velocity (which follows a mixed scaling) 
does not collapse but seems to depend on $Re$ (the collapse is seen in Fig.~\ref{fig:inner_w}).  
% 
While, at a fixed height, the spanwise velocity scales approximately as $Re_\tau^{-1}$, the fraction 
of turning that is encountered within the inner layer of the flow amounts to about 1/3 of the total wind veer 
(Fig.~\ref{fig:etling}a).
% 
The hodographs show the well-known Ekman shape with the laminar profile as an outer limit and 'flatter' 
hodographs, corresponding to less turning, for increasing Reynolds number. 
% 
Importantly, our results suggest that there is no lower limit of the turning. 
% 
This implies that, despite its very large scale separation and huge Reynolds number, the ABL cannot be 
in the 'limit' of high Reynolds number from the perspective of wind veer.

%\subsection{Applications \& Outlook} 
%\begin{itemize}
%\item reference-shear for neutral profile approaches(systematic!) $\rightarrow$ wind engineering! 
%\item initial condition for LES/DNS to eliminate/minimize inertial oscillation in Benchmark simulations
%\item
%\end{itemize}

@@

\vspace{1em}
\noindent\textbf{Acknowledgements} This study is funded by the European Commission's 7$^{th}$ Framework Programme through 
the ERC 2019 Starting Grant trainABL (ID 851374). Simulations were carried out on the German supercomputing systems 
\texttt{hawk} at H\"ocstleistungsrechenzentrum Stuttgart (Bundesprojekt 44187) and \texttt{juwels} at Gauss Centre for supercomputing, 
For\-schungs\-zen\-trum J\"ulich (Projects \texttt{hku24} and \texttt{stadit}). 
We extend our thanks to Sally Issa for valuable feedback on the manuscript.

\section*{Declarations}

\noindent\textbf{Author Contributions} CA conceived the study, acquired funding, performed the simulation, analysed and plotted the data for this work. CA and HW interpreted the data and compiled the manuscript. 
\vspace{1em}

\noindent\textbf{Funding} CA is funded by ERC Starting Grant \emph{trainABL} (ID 851374). Open Access funding for this publication through the DEAL agreement. 
\vspace{1em}

\noindent\textbf{Data Availability} The data is available in long-term repositories and indexed by DOIs as cited in the list of references. Further analysis tools and processed data can be made available upon direct request. \vspace{1em}

\noindent\textbf{Conflict of Interest} The Authors declare no competing interest.
\vspace{1em}

\noindent\textbf{Open Access} This article is licensed under a Creative Commons Attribution 4.0 International License, which permits use, sharing, adaptation, distribution and reproduction in any medium or format, as long as you give appropriate credit to the original author(s) and the source, provide a link to the Creative Commons licence, and indicate if changes were made.  To view a copy of this licence, visit \href{http://creativecommons.org/licenses/by/4.0/}{https://creativecommons.org/licenses/by/4.0/}.

\end{document}
\appendix
\section{Old Stuff}
\todo{OLD FROM HERE} 
Below this region, the gradients in span-wise velocity are rather small and the span-wise velocity
monotonically approaches its surface boundary condition $V(z=0)=0$.
%
While the streamwise velocity follows a universal inner scaling that has acquired its universal, $Re$-independent shape for $Re_D> \mathcal{O}\left(10^3\right)$, the span-wise component that defines how the velocity vector veers when the surface is approached,
does not collapse in inner units, and there is, most importantly no sign of convergence even at the highest Reynolds numbers for
which simulations where carried out.
%
Even though the simplest assumption $V=0$ is reasonable for the lower part of the surface surface layer ($z^-<10^{-3}$), it does
not appropriately capture the profile in the rest of the surface layer:
